Type cones of $\b{g}$-vector fans

### 2019-03-04 thirdPaper/draft/typeConeAssociahedra.tex
minor changes
modifs sect 2.3

%% SPECIFIC NON-KISSING COMPLEX
%% COMBINATORICS
%% quiver
%% walks
%\newcommand{\peaks}[1]{\mathsf{peaks}(#1)} % peaks
%\newcommand{\deeps}[1]{\mathsf{deeps}(#1)} % deeps
%\newcommand{\distinguishedWalk}[2]{\mathsf{dw}(#1,#2)} % distinguished walk
%\newcommand{\distinguishedArrows}[2]{\mathsf{da}(#1,#2)} % distinguished arrows
%\newcommand{\distinguishedString}[2]{\mathsf{ds}(#1,#2)} % distinguished string
%\newcommand{\distinguishedSign}[2]{\varepsilon(#1,#2)} % distinguished sign
%% non-kissing complex, lattice, flip graph
%\newcommand{\kn}{\kappa} % kissing number
%\newcommand{\KN}{\textsc{kn}} % Kissing Number
%\newcommandx{\NKC}[1][1=\quiver]{\mathcal{K}_{\mathrm{nk}}(#1)} % non-kissing complex
%\newcommandx{\RNKC}[1][1=\quiver]{\mathcal{C}_{\mathrm{nk}}(#1)} % non-kissing complex
%\newcommandx{\NKL}[1][1=\quiver]{\mathcal{L}_{\mathrm{nk}}(#1)} % non-kissing lattice
%\newcommandx{\NKG}[1][1=\quiver]{\mathcal{G}_{\mathrm{nk}}(#1)} % non-kissing flip graph
%\newcommandx{\NFC}[1][1=\quiver]{\mathcal{C}_{\mathrm{nf}}(#1)} % non-friendly complex
%\newcommand{\peak}{\mathrm{peak}} % bottom dissection
%\newcommand{\deep}{\mathrm{deep}} % top dissection
%\newcommand{\reversed}[1]{#1^{\mathrm{rev}}} % reverse all arrows
%\renewcommand{\top}{\mathrm{top}} % top
%\newcommand{\bottom}{\mathrm{bot}} % bottom
%\newcommand{\walk}{\operatorname{\omega}} % walk (of a curve)
%% accordion and slalom complexes
%\newcommand{\surface}{\mathcal{S}} % an orientable surface
%\newcommand{\dual}{^*} % duality
%\newcommand{\dissection}{\mathrm{D}} % dissection of a marked surface
%\newcommand{\vertices}{\mathcal{V}} % vertices of a dissection
%\newcommand{\edges}{\mathcal{E}} % edges of a dissection
%\newcommand{\faces}{\mathcal{F}} % faces of a dissection
%\renewcommandx{\AC}[1][1=\dissection]{\mathcal{K}_{\mathrm{acc}}(#1)} % accordion complex
%\newcommandx{\RAC}[1][1=\dissection]{\mathcal{C}_{\mathrm{acc}}(#1)} % reduced accordion complex
%\newcommandx{\SC}[1][1=\dissection\dual]{\mathcal{K}_{\mathrm{sla}}(#1)} % slalom complex
%\newcommandx{\RSC}[1][1=\dissection\dual]{\mathcal{C}_{\mathrm{sla}}(#1)} % reduced slalom complex
%\newcommandx{\NCC}[1][1={\dissection, \dissection\dual}]{\mathcal{K}_{\mathrm{nc}}(#1)} % non-crossing complex
%\newcommandx{\RNCC}[1][1={\dissection, \dissection\dual}]{\mathcal{C}_{\mathrm{nc}}(#1)} % non-crossing complex
%\newcommand{\curveof}{\operatorname{\gamma}} % curve (of a walk)
%\newcommand{\edgeof}{\operatorname{\varepsilon}} % arc of a vertex
%\newcommand{\dualedgeof}{\operatorname{\varepsilon}\dual} % dual arc of a vertex
%\DeclareRobustCommand{\SSS}{\reflectbox{$\mathsf{Z}$}} % S
%\DeclareRobustCommand{\ZZZ}{\mathsf{Z}} % Z
%\newcommand{\vnext}[1]{#1_{\operatorname{next}}} % next point on the boundary of the surface
%\newcommand{\vprevious}[1]{#1_{\operatorname{prev}}} % previous point on the boundary of the surface
%% lattices
%\newcommand{\meet}{\wedge} % meet
%\newcommand{\join}{\vee} % join
%\newcommand{\bigMeet}{\bigwedge} % meet
%\newcommand{\bigJoin}{\bigvee} % join
%\newcommand{\closure}[1]{#1^{\mathrm{cl}}} % closure operator
%\newcommand{\coclosure}[1]{#1^{\mathrm{cocl}}} % coclosure operator
%\newcommand{\uclosure}[1]{#1^{\underline{\mathrm{cl}}}} % loop free closure operator
%\newcommand{\Bicl}[1]{\mathsf{Bic}(#1)} % biclosed sets
%\newcommand{\projDown}{\pi_\downarrow} % down projection map
%\newcommand{\projUp}{\pi^\uparrow} % up projection map
%\newcommand{\JI}{\mathsf{JI}} % join-irreducible
%\newcommand{\MI}{\mathsf{MI}} % meet-irreducible
%\newcommand{\Cong}{\mathsf{Cong}} % congruence lattice
%\newcommand{\con}{\mathrm{con}} % congruence contracting a cover relation
%\newcommand{\ji}{\mathsf{ji}} % join irreducible
%\newcommand{\mi}{\mathsf{mi}} % meet irreducible
%% GEOMETRY
%% d-vectors
%\newcommand{\dvector}[1]{\mathbf{d}(#1)} % d-vector of the path #1
%\newcommand{\dvectors}[1]{\mathbf{d}(#1)} % d-vectors of the paths #1
%\newcommandx{\dvectorFan}[1][1=\quiver]{\mathcal{F}^\mathbf{d}(#1)} % d-vector fan
%% g-vectors
%\newcommand{\gvector}[1]{\mathbf{g}(#1)} % g-vector of the path #1
%\newcommand{\gvectors}[1]{\mathbf{g}(#1)} % g-vectors of the paths #1
%\newcommandx{\gvectorFan}[1][1=\quiver]{\mathcal{F}^\mathbf{g}(#1)} % g-vector fan
%% c-vectors
%\newcommand{\cvector}[2]{\mathbf{c}(#1 \in #2)} % c-vector of the path #1 in the facet #2
%\newcommand{\cvectors}[1]{\mathbf{c}(#1)} % c-vectors of the paths #1
%\newcommandx{\allcvectors}[1][1=\quiver]{\mathbf{C}(#1)} % all c-vectors with respect to the initial cluster #1
%\newcommandx{\cvectorFan}[1][1=\quiver]{\mathcal{F}^\mathbf{c}(#1)} % fan of hyperplanes orthogonal to all c-vectors
%% points, hyperplanes, half-spaces
%\newcommand{\point}[1]{\mathbf{p}(#1)} % vertex of the grid associahedron corresponding to the cluster #1
%\newcommandx{\ray}[1][1=r]{\mathbf{#1}} % ray
%\newcommandx{\rays}[1][1=R]{\mathbf{#1}} % rays
%\newcommand{\hs}{\mathbf{H}^{\le}} % half space
%\newcommand{\HS}[1]{\mathbf{H}^{\le}(#1)} % half space
%\newcommand{\hyp}{\mathbf{H}^{=}} % hyperplane
%\newcommand{\Hyp}[1]{\mathbf{H}^{=}(#1)} % hyperplane
%\newcommand{\fix}[1]{\mathrm{Fix}(#1)} % fix space
%% ALGEBRA
%\newcommand{\Hom}[1]{\operatorname{{\rm Hom}}_{#1}}
%\newcommand{\Ext}[1]{\operatorname{{\rm Ext}}_{#1}}
%\newcommandx{\AR}[1][1=\quiver]{\mathrm{AR}(#1)} % Auslander-Reiten quiver
%\newcommandx{\tTC}[1][1=\quiver]{\mathcal{K}^{\textrm{s$\tau$-tilt}}(#1)} % tau-tilting complex
%\newcommand{\rep}{\operatorname{{\rm rep}}}
%\newcommand{\proj}{\operatorname{{\rm proj}}}
%\newcommand{\koszul}{^!} % duality

%% formating the part command
%\makeatletter
%\def\part{\@startsection{part}{1}%
%\z@{.7\linespacing\@plus\linespacing}{.8\linespacing}%
%{\LARGE\sffamily\centering}}
%\@addtoreset{section}{part}
%\makeatother
%\renewcommand{\thesection}{\arabic{part}.\arabic{section}}

@@

\begin{definition}
The \defn{type cone} of~$\Fan$ is the cone~$\typeCone(\Fan)$ of all $\Fan$-admissible height vectors~$\b{h}$:
\begin{align*}
\typeCone(\Fan) & \eqdef \set{\b{h} \in \R^N}{\Fan \text{ is the normal fan of } P_\b{h}} \\
& = \Bigset{\b{h} \in \R^N}{\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} > 0 \text{ for any adjacent maximal cones~$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan$}}.
\end{align*}
\end{definition}

@@

When~$\Fan$ has the unique exchange relation property, we change the notation~$\coefficient[{\ray[s]}][\rays][\rays']$ to~$\coefficient$ and we obtain that the type cone of~$\Fan$ is expressed as
\[
\typeCone(\Fan) = \Bigset{\b{h} \in \R^N}{\sum_{\ray[s] \in \rays \cup \rays'} \coefficient \, \b{h}_{\ray[s]} > 0 \text{ for any exchangeable rays~$\ray, \ray'$ of~$\Fan$}}.
\]

@@

\para{Cluster algebras}
%
Let~$\Q(x_1, \dots, x_n, p_1, \dots, p_m)$ be the field of rational expressions in~$n+m$ variables with rational coefficients, and~$\Trop{m}$ denote its abelian multiplicative subgroup generated by~$\{p_i\}_{i \in [m]}$.
For~$p = \prod\limits_{i \in [m]} p_i^{a_i} \in \Trop{m}$, define
\(
\positiveExponents{p} \eqdef \prod\limits_{i \in [m]} p_i^{\max(a_i,0)}
\)
and
\(
\negativeExponents{p} \eqdef \prod\limits_{i \in [m]} p_i^{-\min(a_i,0)}
\).
%so that~$p = \positiveExponents{p} \negativeExponents{p}^{-1}$.

@@

Note that \cref{lem:linearDependencegvectorsCA} implies \cref{thm:generalizedAsso} since it ensures that the type cone~$\typeCone(\gvectorFan[\B_\circ])$ of the $\b{g}$-vector fan contains the cone of~${\b{h} \in \R^{\principalVariables}}$ such that
\(
\b{h}_x + \b{h}_{x'} > \sum_{{y \in \cluster \cap \cluster', \; b_{xy} < 0}} -b_{xy}
\)
and
\(
% \quad\text{and}\quad
\b{h}_x + \b{h}_{x'} > \sum_{{y \in \cluster \cap \cluster', \; b_{xy} > 0}} b_{xy} \, \b{h}_y
\)
for any seeds~$(\B, \cluster)$ and~$(\B', \cluster')$ with~${\cluster \ssm \{x\} = \cluster' \ssm \{x'\}}$.

@@

\begin{lemma}
\label{lem:dependenceMesh}
Consider two adjacent seeds~${(\B, \cluster)}$ and~${(\B', \cluster')}$ with~$\cluster \ssm \{x\} = \cluster' \ssm \{x'\}$ such that the entries~$b_{xy}$ for~$y \in \cluster$ are all non-positive or all non-negative.
If either~$x$ or~$x'$ is an initial cluster variable, then
\[
\gvectorFull{\B_\circ}{x} + \gvectorFull{\B_\circ}{x'} = 0.
\]
Otherwise, the $\b{g}$-vectors of~$\cluster \cup \cluster'$ with respect to~$\B_\circ$ satisfy the linear dependence
\[
\gvectorFull{\B_\circ}{x} + \gvectorFull{\B_\circ}{x'} = \sum_{y \in \cluster \cap \cluster'} |b_{xy}| \, \gvectorFull{\B_\circ}{y}.
\]
%\[
%\b{n}(\B_\circ, x, x') = \b{f}_x + \b{f}_{x'} - \sum_{y \in \cluster \cap \cluster'} |b_{xy}| \, \b{f}_y
%\]
\end{lemma}

@@

\subsubsection{Non-kissing complex and non-kissing fan of a gentle quiver}

@@

To define the other family of vectors, we need to recall the notion of distinguished substring of a walk defined in~\cite[Def.~2.25]{PaluPilaudPlamondon-nonkissing}.
Consider an arrow~$\alpha \in Q_1$ contained in two distinct walks~$\omega, \omega'$ of a non-kissing facet~$F \in \NKC$, and let~$\sigma$ denote the common substring of~$\omega$ and~$\omega'$.
We write~$\omega \prec_\alpha \omega'$ if $\omega$ enters and/or exits~$\sigma$ with arrows in the direction pointed by~$\alpha$, while $\omega'$ enters and/or exits~$\sigma$ with arrows in the direction opposite to~$\alpha$.
The \defn{distinguished walk} of the non-kissing facet~$F \in \NKC$ at the arrow~$\alpha$ is the maximum walk of~$F$ for~$\prec_\alpha$.
The \defn{distinguished arrows} of the walk~$\omega$ in the non-kissing facet~$F \in \NKC$ are the arrow where~$\omega$ is the distinguished walk.
It is shown in~\cite[Prop.~2.28]{PaluPilaudPlamondon-nonkissing} that each bending walk in a non-kissing facet has precisely two distinguished arrows, pointing in opposite directions.
The \defn{distinguished substring}~$\distinguishedString{\omega}{F}$ of the walk~$\omega$ in the non-kissing facet~$F \in \NKC$ is the substring located between its two distinguished arrows.

@@

\begin{theorem}[{\cite[Thm.~4.17]{PaluPilaudPlamondon-nonkissing}}]
\label{thm:nonkissingAsso}
For any non-kissing finite gentle quiver~$\quiver$, the set of cones
\[
\gvectorFan[\quiver] \eqdef \set{\R_{\ge 0} \, \gvectors{F}}{F \text{ non-kissing face of } \NKC}
\]
is a complete simplicial fan of~$\R^{Q_0}$, called \defn{non-kissing fan} of~$\quiver$, which realizes the non-kissing complex~$\NKC$.
\end{theorem}

@@

\begin{theorem}[{\cite[Thm.~4.27]{PaluPilaudPlamondon-nonkissing}}]
For any non-kissing finite gentle quiver~$\quiver$, the non-kissing fan~$\gvectorFan[\quiver]$ is the normal fan of the \defn{gentle associahedron}~$\Asso[\quiver]$ defined equivalently as:
\begin{enumerate}[(i)]
\item the convex hull of the points~$\sum_{\omega \in F} \KN(\omega) \, \cvector{F}{\omega}$ for all facets~${F \in \NKC}$,~or
\item the intersection of the halfspaces~$\bigset{\b{x} \in \R^{Q_0}}{\dotprod{\gvector{\omega}}{\b{x}} \le \KN(\omega)}$ for all walks~$\omega$ on~$\bar Q$.
\end{enumerate}
\end{theorem}

@@

In \cref{prop:exchangeablePairsNKC}, the string~$\sigma$ is called the \defn{distinguished} substring of~$\omega$ and~$\omega'$.
We say that a string~$\sigma$ is \defn{distinguishable} if it is the distinguished string of an exchangeable pair.
Note that an equivalent definition of distinguishable strings is given in~\cite[Def.~2.30]{PaluPilaudPlamondon-nonkissing}.
We denote by~$\distinguishableStrings$ the set of distinguishable strings of~$\quiver$.
\vincent{Clean that}

\begin{remark}
\label{rem:exchangeablePairsNKC}
In view of \cref{prop:exchangeablePairsNKC}, it is tempting to look for a characterization of the exchangeable pairs~$\omega, \omega'$ using the kisses between~$\omega$ and~$\omega'$.
However, as illustrated in \cref{fig:nonkissingFans}\,(right), note that
\begin{itemize}
\item two exchangeable walks may kiss along more than one string, but only one is the distinguished string,
\item two non-exchangeable walks can kiss along more than one distinguishable string,
\item two walks that kiss along a single distinguishable string are not always exchangeable,
\item not all strings are distinguishable.
\end{itemize}
\vincent{More explicit. TODO}
\vincent{Do we still have these problems when the gentle algebra is brick and/or without oriented $2$-cycle?}
\end{remark}

\begin{corollary}
For any non-kissing finite gentle quiver~$\quiver$, the type cone of the non-kissing fan~$\gvectorFan[\quiver]$ is given by
\[
\typeCone(\gvectorFan[\quiver]) = \set{\b{h} \in \R^{\walks}}{\begin{array}{l} \b{h}_\omega = 0 \text{ for any improper walk } \omega \\ \b{h}_\omega + \b{h}_{\omega'} > \b{h}_\mu + \b{h}_\nu \text{ for any exchangeable walks } \omega, \omega' \end{array}}.
\]
\end{corollary}

@@

\para{Two families of examples: grid and dissection quivers}
%
We conclude this recollections on non-kissing complexes by two families of examples.
\vincent{TODO}

\bigskip
We are now ready to describe the type cone of the non-kissing fan~$\gvectorFan[\quiver]$.
We first treat the case when~$\quiver$ has no self-kissing walks, which includes both grid and dissection quivers, thus in particular the case of the classical associahedron.
We show that the type cone is then simplicial which gives another proof of \cref{thm:simplicialTypeConeAsso} and \cref{coro:descriptionAllAsso}.

\subsubsection{Simplicial type cones for self-kissing free quivers}

In this section, we focus on the following family of gentle quivers, which was also considered in~\cite[Sect.~4]{}.

@@

The quivers of this family are particularly well-behaved: they avoid all pathologies of \cref{rem:exchangeablePairsNKC} and we will prove that the type cone of their non-kissing fan happens to be simplicial.\pg{Need to exclude two-cycles, else false.}
Note that this family already contains a lot of relevant examples, including:
\begin{itemize}
\item classical associahedra
\item grid associahedra
\item dissection associahedra
\end{itemize}
\vincent{More explicit. TODO}

%For a string~$\sigma$ of~$\quiver$, we denote by $\sigma\hR$ (resp.~$\hL\sigma$, resp.~$\sigma\cR$, resp.~$\cL\sigma$) the unique string of~$\quiver\blossom$ of the form $\sigma\hR = \sigma \alpha_1^{-1} \alpha_2 \dots \alpha_\ell$ (resp.~$\hL\sigma = \alpha_\ell^{-1} \dots \alpha_2^{-1} \alpha_1 \sigma$, resp.~$\sigma\cR = \sigma \alpha_1 \alpha_2^{-1} \dots \alpha_\ell^{-1}$, resp.~$\cL\sigma = \alpha_\ell \dots \alpha_2 \alpha_1^{-1} \sigma$) with~$\ell \ge 1$ and~$\alpha_1, \dots, \alpha_\ell \in Q_1$ and such that~$t(\alpha_\ell)$ (resp.~$s(\alpha_\ell)$, resp.~$s(\alpha_\ell)$, resp.~$t(\alpha_\ell)$) is a blossom of~$\quiver\blossom$.
For a string~$\sigma$ of~$\quiver$, we denote by $\sigma\hR$ (resp.~$\sigma\cR$) the unique string of the blossoming quiver~$\quiver\blossom$ of the form $\sigma\hR = \sigma \alpha_1^{-1} \alpha_2 \dots \alpha_\ell$ (resp.~$\sigma\cR = \sigma \alpha_1 \alpha_2^{-1} \dots \alpha_\ell^{-1}$) with~$\ell \ge 1$ and~${\alpha_1, \dots, \alpha_\ell \in Q_1}$ and such that~$t(\alpha_\ell)$ (resp.~$s(\alpha_\ell)$) is a blossom of~$\quiver\blossom$.
These notations are motivated by the representation of strings used in~\cite{ButlerRingel, PaluPilaudPlamondon-nonkissing}, and the terminology usually says that~$\sigma\hR$ (resp.~$\sigma\cR$) is obtained by adding a \defn{hook} (resp.~\defn{cohook}) to~$\sigma$.
We define similarly~$\hL\sigma$ (resp.~$\cL\sigma$).
The walk~$\hL(\sigma\hR) = (\hL\sigma)\hR$ of~$\quiver$ is simply be denoted~$\hh{\sigma}$, and we define similarly~$\cc{\sigma}$, $\hc{\sigma}$ and~$\ch{\sigma}$.

>>>>

@@

\begin{definition}
The \defn{type cone} of~$\Fan$ is the cone~$\typeCone(\Fan)$ of all $\Fan$-admissible height vectors~$\b{h}$:
\begin{align*}
\typeCone(\Fan) & \eqdef \set{\b{h} \in \R^N}{\Fan \text{ is the normal fan of } P_\b{h}} \\
& = \Bigset{\b{h} \in \R^N}{\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} > 0 \; \begin{array}{l} \text{for any adjacent maximal} \\ \text{cones~$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan$} \end{array}}.
\end{align*}
\end{definition}

@@

When~$\Fan$ has the unique exchange relation property, we change the notation~$\coefficient[{\ray[s]}][\rays][\rays']$ to~$\coefficient$ and we obtain that the type cone of~$\Fan$ is expressed as
\[
\typeCone(\Fan) = \Bigset{\b{h} \in \R^N}{\sum_{\ray[s] \in \rays \cup \rays'} \coefficient \, \b{h}_{\ray[s]} > 0 \text{ for any exchangeable rays~$\ray$ and~$\ray'$ of~$\Fan$}}.
\]

@@

\para{Cluster algebras}
%
Let~$\Q(x_1, \dots, x_n, p_1, \dots, p_m)$ be the field of rational expressions in~$n+m$ variables with rational coefficients, and~$\Trop{m}$ denote its abelian multiplicative subgroup generated by~$\{p_i\}_{i \in [m]}$.
For~$p = \prod\limits_{i \in [m]} p_i^{a_i} \in \Trop{m}$, define
\(
\positiveExponents{p} \eqdef \prod\limits_{i \in [m]} p_i^{\max(a_i,0)}
\)
and
\(
\negativeExponents{p} \eqdef \prod\limits_{i \in [m]} p_i^{-\min(a_i,0)}
\).

@@

Note that \cref{lem:linearDependencegvectorsCA} implies \cref{thm:generalizedAsso} since it ensures that the type cone~$\typeCone(\gvectorFan[\B_\circ])$ of the $\b{g}$-vector fan contains the cone of~${\b{h} \in \R^{\principalVariables}}$ such that
\(
\b{h}_x + \b{h}_{x'} > \sum_{{y \in \cluster \cap \cluster', \; b_{xy} < 0}} -b_{xy}
\)
and
\(
\b{h}_x + \b{h}_{x'} > \sum_{{y \in \cluster \cap \cluster', \; b_{xy} > 0}} b_{xy} \, \b{h}_y
\)
for any seeds~$(\B, \cluster)$ and~$(\B', \cluster')$ with~${\cluster \ssm \{x\} = \cluster' \ssm \{x'\}}$.

@@

\begin{lemma}
\label{lem:dependenceMesh}
Consider two adjacent seeds~${(\B, \cluster)}$ and~${(\B', \cluster')}$ with~$\cluster \ssm \{x\} = \cluster' \ssm \{x'\}$ such that the entries~$b_{xy}$ for~$y \in \cluster$ are all non-positive or all non-negative.
If either~$x$ or~$x'$ is an initial cluster variable, then
\[
\gvectorFull{\B_\circ}{x} + \gvectorFull{\B_\circ}{x'} = 0.
\]
Otherwise, the $\b{g}$-vectors of~$\cluster \cup \cluster'$ with respect to~$\B_\circ$ satisfy the linear dependence
\[
\gvectorFull{\B_\circ}{x} + \gvectorFull{\B_\circ}{x'} = \sum_{y \in \cluster \cap \cluster'} |b_{xy}| \, \gvectorFull{\B_\circ}{y}.
\]
\end{lemma}

@@

\subsubsection{Non-kissing complex and non-kissing fan of a gentle quiver}
\label{subsubsec:nonkissingComplex}

@@

To define the other family of vectors, we need to recall the notion of distinguished substring of a walk defined in~\cite[Def.~2.25]{PaluPilaudPlamondon-nonkissing}.
Consider an arrow~$\alpha \in Q_1$ contained in two distinct walks~$\omega, \omega'$ of a non-kissing facet~$F \in \NKC$, and let~$\sigma$ denote the common substring of~$\omega$ and~$\omega'$.
We write~$\omega \prec_\alpha \omega'$ if $\omega$ enters and/or exits~$\sigma$ with arrows in the direction pointed by~$\alpha$, while $\omega'$ enters and/or exits~$\sigma$ with arrows in the direction opposite to~$\alpha$.
The \defn{distinguished walk} of the non-kissing facet~$F \in \NKC$ at the arrow~$\alpha$ is the maximum walk of~$F$ for~$\prec_\alpha$.
The \defn{distinguished arrows} of the walk~$\omega$ in the non-kissing facet~$F \in \NKC$ are the arrow where~$\omega$ is the distinguished walk.
It is shown in~\cite[Prop.~2.28]{PaluPilaudPlamondon-nonkissing} that each bending walk in a non-kissing facet has precisely two distinguished arrows, pointing in opposite directions.
The \defn{distinguished substring}~$\distinguishedString{\omega}{F}$ of the walk~$\omega$ in the non-kissing facet~$F \in \NKC$ is the substring located between its two distinguished arrows.
We say that a string~$\sigma$ is \defn{distinguishable} if it is the distinguished string of a walk in a non-kissing facet, and we denote by~$\distinguishableStrings$ the set of distinguishable strings of~$\quiver$.

@@

\begin{theorem}[{\cite[Thm.~4.17]{PaluPilaudPlamondon-nonkissing}}]
\label{thm:nonkissingFan}
For any non-kissing finite gentle quiver~$\quiver$, the set of cones
\[
\gvectorFan[\quiver] \eqdef \set{\R_{\ge 0} \, \gvectors{F}}{F \text{ non-kissing face of } \NKC}
\]
is a complete simplicial fan of~$\R^{Q_0}$, called \defn{non-kissing fan} of~$\quiver$, which realizes the non-kissing complex~$\NKC$.
\end{theorem}

@@

\begin{theorem}[{\cite[Thm.~4.27]{PaluPilaudPlamondon-nonkissing}}]
\label{thm:nonkissingAsso}
For any non-kissing finite gentle quiver~$\quiver$, the non-kissing fan~$\gvectorFan[\quiver]$ is the normal fan of the \defn{gentle associahedron}~$\Asso[\quiver]$ defined equivalently as:
\begin{enumerate}[(i)]
\item the convex hull of the points~$\sum_{\omega \in F} \KN(\omega) \, \cvector{F}{\omega}$ for all facets~${F \in \NKC}$,~or
\item the intersection of the halfspaces~$\bigset{\b{x} \in \R^{Q_0}}{\dotprod{\gvector{\omega}}{\b{x}} \le \KN(\omega)}$ for all walks~$\omega$ on~$\bar Q$.
\end{enumerate}
\end{theorem}

@@

\begin{remark}
\label{rem:exchangeablePairsNKC}
In view of \cref{prop:exchangeablePairsNKC}, it is tempting to look for a characterization of the exchangeable pairs~$\omega, \omega'$ using the kisses between~$\omega$ and~$\omega'$.
This question was discussed in \cite[Sect.~9]{BrustleDouvilleMousavandThomasYildirim}.
However, as illustrated in \cref{fig:nonkissingFans}\,(right), note that
\begin{itemize}
\item two exchangeable walks may kiss along more than one string (only one is distinguished),
\item two non-exchangeable walks can kiss along more than one distinguishable string,
\item two walks that kiss along a single distinguishable string are not always exchangeable,
\item not all strings are distinguishable.
\end{itemize}
In \cref{subsubsec:simplicialTypeConeNKC}, we will restrict to a situation that avoids all these patologies.
\vincent{Do we still have these problems when the gentle algebra is brick and/or without oriented $2$-cycle?}
\end{remark}

@@

\para{Two families of examples: grid and dissection quivers}
%
We conclude this recollections on non-kissing complexes by two families of examples:
\begin{enumerate}[(1)]
\item \textbf{Grid quivers}: Consider the infinite grid quiver~$\quiver_{Z^2}$, whose vertices are all integer points of~$\Z^2$, whose arrows connect between~$(i,j) \longrightarrow (i,j+1)$ and~$(i,j) \longrightarrow (i+1,j)$ for any~$(i,j) \in \Z^2$, and whose relations forbid~$(i-1,j) \longrightarrow (i,j) \longrightarrow (i,j+1)$ and~$(i,j-1) \longrightarrow (i,j) \longrightarrow (i+1,j)$ for any~$(i,j) \in \Z^2$. A \defn{gentle grid quiver} is any subquiver~$\quiver_A$ of~$\quiver_{Z^2}$ induced by a finite subset~$A \subset \Z^2$ of the integer grid. The non-kissing complex of gentle grid quivers were introduced in~\cite{McConville} with motivation coming from~\cite{PetersenPylyavskyySpeyer} and~\cite{SantosStumpWelker}.

\smallskip
\item \textbf{Dissection quivers}: Consider a dissection~$D$ of a convex polygon~$P$ (that is a crossing-free set of diagonals of~$P$) and its \defn{gentle dissection quiver}~$\quiver_D$, whose vertices are the internal diagonals of~$D$, whose arrows connect pairs of consecutive internal diagonals along the boundary of a face of~$D$, and whose relations correspond to triples of consecutive internal diagonals along the boundary of a face of~$D$. The non-kissing complex of~$\quiver_D$ the corresponds to non-crossing sets of accordions of~$D$, where an accordion is a segment connecting the middles of two boundary edges of~$P$ and crossing a connected set of diagonals of~$D$. This accordion complex was studied in~\cite{GarverMcConville} and~\cite{MannevillePilaud-accordion} with motivation coming from~\cite{Baryshnikov}, and~\cite{Chapoton-quadrangulations}.
\end{enumerate}

\vincent{Example picture?}

\smallskip
These two families of non-kissing complexes are well-behaved as they avoid all pathologies of \cref{rem:exchangeablePairsNKC}.
However, they still provide good examples of non-kissing complexes.
In particular, both families contain the classical associahedron.
Namely, the gentle associahedron~$\Asso[\quiver]$ is the classical associahedron of~\cite{ShniderSternberg, Loday} presented in \cref{thm:associahedronLoday} when~${\quiver = \quiver_A}$ for the path~${A = \set{(0,j)}{j \in [n]}}$ or equivalently~$\quiver = \quiver_D$ for the fan triangulation~$D$ (where all internal diagonals incident to the same point).
More generally, $\Asso[\quiver]$ is an associahedron of~\cite{HohlwegLange} when~$\quiver = \quiver_A$ for a ribbon~$A$ or equivalently~$\quiver = \quiver_D$ for a triangulation~$D$.

Note that, it was shown in~\cite{PaluPilaudPlamondon-surfaces} that the accordion complexes can be extended to dissections of arbitrary orientable surfaces with marked points and then provide a geometric model for all non-kissing complexes of gentle quivers.

\subsubsection{Type cones of non-kissing fans}

We now discuss the type cones of the non-kissing fans defined in \cref{thm:nonkissingFan}. We obtain from \cref{prop:exchangeablePairsNKC} the following redundant description.

\begin{corollary}
For any non-kissing finite gentle quiver~$\quiver$, the type cone of the non-kissing fan~$\gvectorFan[\quiver]$ is given by
\[
\typeCone(\gvectorFan[\quiver]) = \set{\b{h} \in \R^{\walks}}{\begin{array}{l} \b{h}_\omega = 0 \text{ for any improper walk } \omega \\ \b{h}_\omega + \b{h}_{\omega'} > \b{h}_\mu + \b{h}_\nu \text{ for any exchangeable walks } \omega, \omega' \end{array}},
\]
where the walks~$\mu$ and~$\nu$ for two exchangeable walks~$\omega, \omega'$ are defined in \cref{prop:exchangeablePairsNKC}.
\end{corollary}

\begin{example}
\label{exm:typeConeNKC}
Consider the gentle quivers of \cref{fig:typeConeNKC}.
\vincent{Todo}
\end{example}

As illustrated in \cref{exm:typeConeNKC}, the type cone of the non-kissing fan is not simplicial and we do not always understand its extremal exchangeable pairs.
In the next section section, we will explore a special family of gentle quivers for which we can completely describe the type cone.
The combination of this special family with computer experiments in the general case support the following conjecture.

\begin{conjecture}
Consider a distinguishable string~$\sigma \in \distinguishableStrings$, and let~$\omega$ (resp.~$\omega'$) be the walk obtained from~$\sigma$ by adding two hooks (resp.~two cohooks) at the endpoints of~$\sigma$. Then
\begin{enumerate}
\item The $\b{c}$-vector~$\multiplicityVector_\sigma$ is the direction of at least one extremal exchangeable pair.
\item If the walks~$\omega$ and~$\omega'$ are non-self-kissing and exchangeable, then they form the unique extremal exchangeable pair directed by~$\sigma$. These extremal exchangeable pairs correspond to meshes of the Auslander-Reiten quiver.
\end{enumerate}
\end{conjecture}

\subsubsection{Simplicial type cones for self-kissing-free and $2$-cycle-free quivers}
\label{subsubsec:simplicialTypeConeNKC}

In this section, we focus on the following family of gentle quivers, which was also considered in~\cite[Sect.~4]{GarverMcConvilleMousavand}.

@@

We say that a gentle quiver~$\quiver$ is \defn{simple} when
\begin{itemize}
\item it satisfies the three equivalent conditions of \cref{prop:noSelfKissing},
\item it contains no cycle of length~$2$.
\end{itemize}
Note that this family already contains a lot of relevant examples, including the grid and dissection quivers discussed in \cref{subsubsec:nonkissingComplex}.
In particular, the classical associahedron is the gentle associahedron of a simple gentle quiver.
We will see in~\cref{coro:simplicialTypeConeNKC} that the type cone of the non-kissing fan~$\gvectorFan[\quiver]$ of a simple gentle quiver~$\quiver$ happens to be simplicial, and we will derive in~\cref{coro:allPolytopalRealizationsNKC} a simple description of all polytopal realizations of~$\gvectorFan[\quiver]$.

For a string~$\sigma$ of~$\quiver$, we denote by $\sigma\hR$ (resp.~$\sigma\cR$) the unique string of the blossoming quiver~$\quiver\blossom$ of the form $\sigma\hR = \sigma \alpha_1^{-1} \alpha_2 \dots \alpha_\ell$ (resp.~$\sigma\cR = \sigma \alpha_1 \alpha_2^{-1} \dots \alpha_\ell^{-1}$) with~$\ell \ge 1$ and~${\alpha_1, \dots, \alpha_\ell \in Q_1}$ and such that~$t(\alpha_\ell)$ (resp.~$s(\alpha_\ell)$) is a blossom of~$\quiver\blossom$.
These notations are motivated by the representation of strings used in~\cite{ButlerRingel, PaluPilaudPlamondon-nonkissing}, and the terminology usually says that~$\sigma\hR$ (resp.~$\sigma\cR$) is obtained by adding a \defn{hook} (resp.~\defn{cohook}) to~$\sigma$.
We define similarly~$\hL\sigma$ (resp.~$\cL\sigma$).
The walk~$\hL(\sigma\hR) = (\hL\sigma)\hR$ of~$\quiver$ is simply be denoted~$\hh{\sigma}$, and we define similarly~$\cc{\sigma}$, $\hc{\sigma}$ and~$\ch{\sigma}$.

\begin{proposition}
For any simple gentle quiver~$\quiver$ and any string~$\sigma \in \strings$, the walks~$\cc{\sigma}$ and~$\hh{\sigma}$ are exchangeable with distinguished substring~$\sigma$.
\end{proposition}
